---
title: "R Notebook"
output: html_notebook
---

Load all libraries required

```{r}
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(knitr)
library(stringr)
```

Read all the happiness index csv files and read them into a list of dataframes
```{r}
mydir = "input"
myfiles_links = list.files(path=mydir, pattern="*.csv", full.names=TRUE)
myfiles <- lapply(myfiles_links, read_delim)
```

Each dataframe has a different set of columns (all are almost the same but with 
slight variation in names of columns). Doing a small tidying job to standardize 
all the column names
```{r}
for (i in 1:5){
  myfiles[[i]]$Year = str_remove(str_remove(myfiles_links[[i]], "input/"), ".csv")
  if ("Happiness Rank" %in% colnames(myfiles[[i]]))
  {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(`Overall.rank` = `Happiness Rank`)
  }
  if ("Happiness.Rank" %in% colnames(myfiles[[i]]))
  {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(`Overall.rank` = `Happiness.Rank`)
  }
  if ("Overall rank" %in% colnames(myfiles[[i]]))
  {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(`Overall.rank` = `Overall rank`)
  }
  if ("Country or region" %in% colnames(myfiles[[i]]))
  {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(Country = `Country or region`)
  }
  if ("Perceptions of corruption" %in% colnames(myfiles[[i]])) {
    myfiles[[i]]$`Perceptions of corruption` = 
      as.double(myfiles[[i]]$`Perceptions of corruption`)
  }
  if ("GDP per capita" %in% colnames(myfiles[[i]]))
  {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(`Economy (GDP per Capita)` = `GDP per capita`)
  }
  if ("Economy..GDP.per.Capita." %in% colnames(myfiles[[i]]))
  {
    print(i)
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(`Economy (GDP per Capita)` = `Economy..GDP.per.Capita.`)
  }
  if ("Trust..Government.Corruption." %in% colnames(myfiles[[i]]))
  {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(`Trust (Government Corruption)` = `Trust..Government.Corruption.`)
  }
  if ("Freedom to make life choices" %in% colnames(myfiles[[i]]))
  {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(Freedom = `Freedom to make life choices`)
  }
  if ("Happiness Score" %in% colnames(myfiles[[i]]))
  {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(Happiness.Score = `Happiness Score`)
  }
  if ("Score" %in% colnames(myfiles[[i]])) {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(Happiness.Score = Score)
  }
  if ("Upper Confidence Interval" %in% colnames(myfiles[[i]])) {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(Whisker.high = `Upper Confidence Interval`)
  }
  if ("Lower Confidence Interval" %in% colnames(myfiles[[i]])) {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(Whisker.low = `Lower Confidence Interval`)
  }
  if ("Healthy life expectancy" %in% colnames(myfiles[[i]])) {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(`Health (Life Expectancy)` = `Healthy life expectancy`)
  }
  if ("Health..Life.Expectancy." %in% colnames(myfiles[[i]])) {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(`Health (Life Expectancy)` = `Health..Life.Expectancy.`)
  }
  if ("Dystopia Residual" %in% colnames(myfiles[[i]])) {
    myfiles[[i]] <- myfiles[[i]] %>%
      rename(`Dystopia.Residual` = `Dystopia Residual`)
  }
}
dat_csv <- dplyr::bind_rows(myfiles)

```
The files in the latter years had the region and country combined together in 
one column (although the value was the country name), hence filling up the 
Region with the Region value of the same country from previous years. 
```{r}
dat_csv <- dat_csv %>% 
  group_by(Country) %>% 
  fill(Region) %>% #default direction down
  fill(Region, .direction = "up")
```


Imputing NA values with the average values and cleaning up the data to make sure 
that all NA values are imputed
```{r}
dataset <- dat_csv %>% 
  group_by(Country) %>% 
  mutate(`Social support`= ifelse(is.na(`Social support`),
                                  mean(`Social support`,na.rm = T),
                                  `Social support`)) %>%
  mutate(`Perceptions of corruption` = ifelse(is.na(`Perceptions of corruption`),
                                              mean(`Perceptions of corruption`, 
                                                   na.rm = T), 
                                              `Perceptions of corruption`)) %>%
  mutate(`Whisker.low` = `Happiness.Score`) %>%
  mutate(`Whisker.high` = `Happiness.Score`) %>%
  mutate(`Trust (Government Corruption)` = 
           ifelse(is.na(`Trust (Government Corruption)`), 
                  median(`Trust (Government Corruption)`, na.rm = T), 
                  `Trust (Government Corruption)`)) %>%
  mutate(`Dystopia.Residual` = ifelse(is.na(`Dystopia.Residual`), 
                                      median(`Dystopia.Residual`, na.rm = T), 
                                      `Dystopia.Residual`)) %>%
  mutate(`Standard Error`= ifelse(is.na(`Standard Error`), 0, 
                                  `Standard Error`)) %>%
  mutate(`Family` = ifelse(is.na(`Family`), 
                           median(`Family`, na.rm = T), `Family`)) %>%
  mutate(Region = ifelse(is.na(Region), "", Region))
  
dataset[is.na(dataset)] <- 0
```

```{r}
colnames(dataset)
```

```{r}
subset(dataset, select = c("Trust (Government Corruption)", "Perceptions of corruption", "Year", "Country"))
```


```{r}
subset(dataset, select = c("Family", "Social support", "Year", "Country"))
```

Distribution of Happiness Scores for each region
```{r}
dataset %>%
  filter(Region != "") %>%
  ggplot(aes(x=as.factor(Region), y=Happiness.Score, color=Year)) +
  geom_boxplot() + 
  coord_flip() +
  labs(title="Distribution of Happiness Scores per Region", x="Region", 
       y="Happiness Score")
```
From the above visualization, we can see that counties in Western Europe and 
Australia-New Zealand, and North America have consistently high happiness 
scores. Similarly, countries in Africa, Southern Asia have lower happiness 
scores. This ties into the logical assumption that countries that are 
more developed have higher happiness scores. Middle East and Northern Africa
have countries that range between both ends of the happiness score spectrum 
which makes sense since some countries in the middle east (Dubai) are very 
developed. 

```{r}
top10countries2015 = dataset %>% 
  filter(Year == 2015) %>%
  filter(Overall.rank <= 10) %>%
  select(Country)

dataset %>%
  filter(Country %in% as.list(top10countries2015$Country)) %>%
  ggplot( aes(x=as.factor(Year), y=Overall.rank, color=as.factor(Country))) +
    geom_line(aes(color=Country, group=Country)) +
  geom_point() +
  labs(title="Top 10 countries with the highest happiness ranks over the years",
       x="Year", y="Rank")


```
The top 10 countries with the highest happiness scores in 2016 are taken in this
visualization to see the progression of ranks over the next 5 years. One thing, 
that we can clearly see is that there is a lot of shuffling between the ranks
of the top 10 countries but it is not very likely that they leave the top 10 
spots. (Only one country that was in the top 10 in 2015 went out of the top 10: 
in 2019, Australia was ranked 11). Clearly the top 10 countries are consistent
leading to happy citizens. 

```{r}
bottom10countries2015 <- dataset %>% 
  filter(Year == 2015) %>%
  arrange(Overall.rank) %>%
  tail(10) %>%
  select(Country)

dataset %>%
  filter(Country %in% as.list(bottom10countries2015$Country)) %>%
  ggplot( aes(x=as.factor(Year), y=Overall.rank, color=as.factor(Country))) +
    geom_line(aes(color=Country, group=Country)) +
  geom_point() +
  labs(title="10 Countries with the least happiness 
       ranks in 2015 over the years", x="Year", y="Rank")

```
The top 10 countries with the least happiness scores in 2016 are taken in this
visualization to see the progression of ranks over the next 5 years. This is 
very rewarding to see that all of the 10 countries increased their ranks over
the next years; some countries even jumped over 50 spots: Ivory Coast. This 
could be due to drastic changes in policy, etc.


```{r}
dataset %>%
  ggplot(aes(x=Happiness.Score, y=`Economy (GDP per Capita)`)) +
  geom_point() + 
  geom_smooth() +
  facet_wrap(.~Year) + 
  labs(title="Happiness Score vs Economy", x="Happiness Score", y="Economy")
```
We can see that there is a positive linear relationship between the Economy and
the Happiness Score for a country over each year. 

```{r}
dataset %>%
  ggplot(aes(x=Happiness.Score, y=Family)) +
  geom_point() + 
  geom_smooth() +
  facet_wrap(.~Year) +
  labs(title="Happiness Score vs Family", x="Happiness Score", y="Family")
```
We can see that there is a positive linear relationship between the Family and
the Happiness Score for a country over each year. 

```{r}
dataset %>%
  ggplot(aes(x=Happiness.Score, y=`Health (Life Expectancy)`)) + 
  geom_point() +
  geom_smooth() +
  facet_wrap(.~Year) + 
  labs(title="Happiness Score vs Life Expectancy", x="Happiness Score", y="
       Life Expectancy")
```
We can see that there is a positive linear relationship between the 
Life Expectancy and the Happiness Score for a country over each year. 

```{r}
dataset %>%
  ggplot(aes(x=Happiness.Score, y=Freedom)) + 
  geom_point() +
  geom_smooth() + 
  facet_wrap(.~Year) +
  labs(title="Happiness Score vs Freedom", x="Happiness Score", y="Freedom") 
  
```
We can see that there is a positive almost-linear relationship between the 
Life Expectancy and the Happiness Score for a country over each year. In some
years (2015, 2018) the relationship is slightly curved. 


```{r}
dataset %>%
  ggplot(aes(x=log2(Happiness.Score), y=Generosity)) + 
  geom_point() +
  geom_smooth() + 
  facet_wrap(.~Year) +
  labs(title="Happiness Score vs Generosity", x="Happiness Score", 
       y="Generosity") 
```
We can see that there is no linear relationship between Generosity 
and the Happiness Score for a country over each year. 

```{r}
dataset %>%
  ggplot(aes(x=Happiness.Score, y=`Trust (Government Corruption)`)) + 
  geom_point() +
  geom_smooth() +
  facet_wrap(.~Year) + 
  labs(title="Happiness Score vs Trust (Government Corruption)", 
       x="Happiness Score")
```
We can see that there is a relationship between the 
Trust in Governement over the years. However, the relationship is not 
linear.


```{r}
glimpse(dataset)
```

```{r}
dataset %>%
  select(Country, Happiness.Score) %>%
  group_by(Country) %>%
  summarise(Avg_Happiness=mean(Happiness.Score, na.rm=TRUE)) %>%
  arrange(desc(Avg_Happiness)) %>%
  slice_head(n = 25) %>%
  ggplot(aes(reorder(Country, Avg_Happiness), Avg_Happiness)) +
  geom_point(color = "lightblue") +
  theme_bw() + 
  coord_flip() +
  labs(title = "Top 25 Happiest Countries based on Average Score for 2015-2019", 
       x = "Country", y = "Happiness Score")
```
What we have here is a list of the top 25 countries based on the average 
happiness score over the years 2015-2019. Denmark is the happiest country 
with Norway a closed second and Finland third.


```{r}
world <- map_data("world")
whr <- dataset %>%
  select(Country, Happiness.Score) %>%
  group_by(Country) %>%
  summarise(Score=mean(Happiness.Score, na.rm=TRUE)) %>%
  mutate(country = str_trim(Country)) %>% 
  mutate(country = case_when(country == 'United States' ~ 'USA',
                                country == 'United Kingdom' ~ 'UK',
                                TRUE ~ country))
setdiff(whr$country, world$region)
map <- left_join(world, whr, by = c('region' = 'country'))
ggplot(data = map, mapping = aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = Score))  +
  scale_fill_continuous(low="thistle2", high="darkred", na.value="snow2") +
  coord_quickmap() +
  labs(title = "Happiness Around the World") +
  theme_void()
```

This is a choropleth of the world showing the average Happiness Score for each
region. As theorized before, North America, Western Europe and Australia seem
to have high Happiness Scores and some countries in Asia, Africa are on the lower
end of the range. 

```{r}
dataset %>% ggplot(aes(x=Happiness.Score, y =`Social support`)) + 
  geom_point() + facet_wrap(~as.factor(Year)) + geom_smooth(lm = loess) + labs(title = "Social Support vs Happiness")

```
The relationship between social support and happiness shows an increasing linear 
curve which has straightened over the years.

```{r}
dataset %>% ggplot(aes(x=Happiness.Score, y =`Perceptions of corruption`)) + 
  geom_point() + facet_wrap(~as.factor(Year)) + geom_smooth(lm = loess) + labs(title = "Perception of corruption vs Happiness")

```
The graph for Happiness vs Perception of Corruption shows an interesting curve 
which has a rounded bottom over the median and has a slightly increasing 
relationship for higher values of Happiness.

The next series of plots are  very interesting correlation matrices between a select 
number of attributes plotted using a heat-map over the years. We can see that 
the the correlation remains the same over the years. There is a high
negative correlation between Happiness Score and Rank (ie) larger your happiness
score, the more smaller your rank (which makes it higher). We can also see that
there is no correlation between generosity and the score which was what we had
previously hypoethesized using the graphs. What's interesting is that Freedom to 
make Life Choices does not have a huge correlation with the Happiness Score
especially compared to other factors such as GDP, etc. 

```{r}
if (!require("corrplot")) install.packages("corrplot")
library(corrplot)
whr2019 <- dataset %>%
  filter(Year == 2019) %>%
  select(Overall.rank, Happiness.Score, `Economy (GDP per Capita)`, 
         `Health (Life Expectancy)`, Freedom, `Social support`, 
         `Perceptions of corruption`)
temp <-sapply(whr2019, is.numeric)
cor_data <- cor(whr2019[,temp])
corrplot(cor_data, method="square", type='full', tl.cex=0.8, tl.col = 'black')
```

```{r}
whr2018 <- dataset %>%
  filter(Year == 2018) %>%
  select(Overall.rank, Happiness.Score, `Economy (GDP per Capita)`, 
         `Health (Life Expectancy)`, Freedom, `Social support`, 
         `Perceptions of corruption`)
temp <-sapply(whr2018, is.numeric)
cor_data <- cor(whr2018[,temp])
corrplot(cor_data, method="square", type='full', tl.cex=0.8, tl.col = 'black')
```
```{r}
whr2017 <- dataset %>%
  filter(Year == 2017) %>%
  select(Overall.rank, Happiness.Score, `Economy (GDP per Capita)`, 
         `Health (Life Expectancy)`, Freedom, `Family`, 
         `Trust (Government Corruption)`)
temp <-sapply(whr2017, is.numeric)
cor_data <- cor(whr2017[,temp])
corrplot(cor_data, method="square", type='full', tl.cex=0.8, tl.col = 'black')
```

```{r}
whr2016 <- dataset %>%
  filter(Year == 2016) %>%
  select(Overall.rank, Happiness.Score, `Economy (GDP per Capita)`, 
         `Health (Life Expectancy)`, Freedom, `Family`, 
         `Trust (Government Corruption)`)
temp <-sapply(whr2016, is.numeric)
cor_data <- cor(whr2016[,temp])
corrplot(cor_data, method="square", type='full', tl.cex=0.8, tl.col = 'black')
```


```{r}
whr2015 <- dataset %>%
  filter(Year == 2015) %>%
  select(Overall.rank, Happiness.Score, `Economy (GDP per Capita)`, 
         `Health (Life Expectancy)`, Freedom, `Family`, 
         `Trust (Government Corruption)`)
temp <-sapply(whr2015, is.numeric)
cor_data <- cor(whr2015[,temp])
corrplot(cor_data, method="square", type='full', tl.cex=0.8, tl.col = 'black')
```

Adding Gapminder dataset to the present dataset to see if there are any 
additional statistical relationships

```{r}
whrgm <- read_csv('happiness_with_gapminder.csv')
glimpse(whrgm)

library(gapminder)
glimpse(gapminder)
```


```{r}
whrgm <- whrgm %>%
  rename(`Employment Rate Percent` = aged_15plus_employment_rate_percent,
         `Labour Force Participation` = aged_15plus_labour_force_participation_rate_percent,
         `Child Mortality per 1000` = child_mortality_0_5_year_olds_dying_per_1000_born,
         `Suicides per 100000` = suicide_per_100000_people,
         `Cell Phones per 100` = cell_phones_per_100_people)
temp <-sapply(whrgm, is.numeric)
cor_data <- cor(whrgm[,temp])
corrplot(cor_data, method="square", type='full', tl.cex=0.8, tl.col = 'black')
```
Checking the correlation matrix with the added gapminder attributes, we can 
first see that there is a good negative correlation between Child mortality
and Happiness Score; the lower the child mortality rate, the higher the happiness
score. What is interesting is that the number of suicides does not have much 
correlation with the happiness score (in fact it is almost positive) which does 
not make sense. The employment rate percent and the labour force participation 
have complete positive correlation with eachother and hence for all intents and
purposes of statistical inference can be taken as one attribute. Another thing 
that is interesting is that Economy has a very high negative correlation with 
Child Mortality. The lower the economy the higher the child mortality. 

```{r}
whrgm %>% ggplot(aes(x=`Happiness Score`, y =`Employment Rate Percent`)) + 
  geom_point() + facet_wrap(~as.factor(Year)) + geom_smooth(lm = loess) + labs(title = "Employement Rate vs Happiness")
```
The relationships between Employment rate and Happiness Score changes over the 
years (2, to be exact). In the first year, it has a weak positive linear 
relationship with Happiness Score. However, in the second year, it is almost a
sine wave where at one point the happiness score is high for a lower employment
rate. This could mean that people don't consider employment to be the sole factor
in gauging their happiness.

```{r}
whrgm %>% ggplot(aes(x=`Happiness Score`, y =`Labour Force Participation`)) + 
  geom_point() + facet_wrap(~as.factor(Year)) + geom_smooth(lm = loess) + labs(title = "Labour Force Participation vs Happiness")
```
As mentioned before, Labour Force and Employment rate have a very strong positive
correlation and we can see that being reflected in their relationships with
the Happiness Score. Similar to the previous visualization, there is a weak 
positive linear relationship in 2015 but in 2016, it is not linear. 

```{r}
whrgm %>% ggplot(aes(x=`Happiness Score`, y =`Suicides per 100000`)) + 
  geom_point() + facet_wrap(~as.factor(Year)) + geom_smooth(lm = loess) + 
  labs(title = "Suicide Rate vs Happiness")
```
Suicide Rate and Happiness Score do not have a linear relationship. We can see 
that in 2015, it starts off slightly linear and then curves and in 2016, there is 
no linearity in the relationship. This is interesting since suicides
are linked with sadness and depression but however does not play a part in 
actually computing the happiness score.